### 2017-03-09 Document/LaTeX/ArticleExample.tex
ENH:  INitial commit.

%
% Complete documentation on the extended LaTeX markup used for Insight
% documentation is available in ``Documenting Insight'', which is part
% of the standard documentation for Insight.  It may be found online
% at:
%
%     http://www.itk.org/

\documentclass{InsightArticle}

\usepackage[dvips]{graphicx}
\usepackage[cmtip,all]{xy}
\newcommand{\longsquiggly}{\xymatrix{{}\ar@{~>}[r]&{}}}
\usepackage{mathtools}
\usepackage{subcaption}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  hyperref should be the last package to be loaded.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage[
bookmarks,
bookmarksopen,
backref,
colorlinks,linkcolor={blue},citecolor={blue},urlcolor={blue},
]{hyperref}

%  This is a template for Papers to the Insight Journal.
%  It is comparable to a technical report format.

% The title should be descriptive enough for people to be able to find
% the relevant document.
\title{Two Luis Miguel fans walk into a bar in Nagoya $\xRightarrow[]{yada\,\,\,yada\,\,\,yada}$
an ITK-implementation of a popular patch-based denoising filter}

%
% NOTE: This is the last number of the "handle" URL that
% The Insight Journal assigns to your paper as part of the
% submission process. Please replace the number "1338" with
% the actual handle number that you get assigned.
%
\newcommand{\IJhandlerIDnumber}{3564}

% Increment the release number whenever significant changes are made.
% The author and/or editor can define 'significant' however they like.
\release{0.00}

% At minimum, give your name and an email address.  You can include a
% snail-mail address if you like.
\author{Nick Tustison$^{1}$ and Jose Vicente Manj\'on$^{2}$}
\authoraddress{$^{1}$University of Virginia, Charlottesville, VA\\
               $^{2}$ITACA institute, Universidad Polit\'ecnica de Valencia, Valencia, Spain
               }

\begin{document}

%
% Add hyperlink to the web location and license of the paper.
% The argument of this command is the handler identifier given
% by the Insight Journal to this paper.
%
\IJhandlefooter{\IJhandlerIDnumber}

%\ifpdf
%\else
%   %
%   % Commands for including Graphics when using latex
%   %
%   \DeclareGraphicsExtensions{.eps,.jpg,.gif,.tiff,.bmp,.png}
%   \DeclareGraphicsRule{.jpg}{eps}{.jpg.bb}{`convert #1 eps:-}
%   \DeclareGraphicsRule{.gif}{eps}{.gif.bb}{`convert #1 eps:-}
%   \DeclareGraphicsRule{.tiff}{eps}{.tiff.bb}{`convert #1 eps:-}
%   \DeclareGraphicsRule{.bmp}{eps}{.bmp.bb}{`convert #1 eps:-}
%   \DeclareGraphicsRule{.png}{eps}{.png.bb}{`convert #1 eps:-}
%\fi

\maketitle

\ifhtml
\chapter*{Front Matter\label{front}}
\fi

% The abstract should be a paragraph or two long, and describe the
% scope of the document.
\begin{abstract}
\noindent
This document describes an ITK implementation of the adaptive patch-based
image denoising algorithm described in \cite{Manjon:2010aa}.  This offering consists
of the templated class itkPatchBasedDenoisingImageFilter interfaced through
the DenoiseImage program which includes its own interface program.  We include
a couple of examples.
\end{abstract}

\IJhandlenote{\IJhandlerIDnumber}

% \tableofcontents

\section{Introduction}

Image filtering techniques are usually applied to increase the quality of MR images.
Most of these techniques assume an equal noise distribution across the image. When this
assumption is not met, the resulting filtering becomes suboptimal. This is the case of MR
images with spatially varying noise levels, such as those obtained by parallel imaging
(sensitivity-encoded), intensity inhomogeneity-corrected images or surface coil based
acquisitions. In this work, we have adapted a recently proposed filter so-called
Non-Local Means (NL-means) to deal with MR images with spatially varying noise levels
(for both Gaussian and Rician distributed noise). With this new method, information
regarding the local image noise level is used to adjust the amount of denoising strength
of the filter. Such information is automatically obtained from the images by using the
local noise estimation method described in \cite{Manjon:2010aa}.

This method has been validated and compared with the standard NL-means filter on simulated
and real MR imaging data. This noise-adaptive NL-means method was demonstrated to outperform
the standard NL-means based filter when spatially varying noise is present in the images.
Although source code is available from Prof. Manj\'on Herrera's website
(http://personales.upv.es/jmanjon/denoising/arnlm.html).  This particular algorithm is
currently not available in the Insight Toolkit which provides good performance with minimal
parameter tuning. In this Insight Journal submission, we provide a brief overview of the ITK
implementation and provide a couple examples.

\section{ITK Implementation}

The following files define the adaptive NL-means algorithm and is written in ITK
coding style:
\begin{itemize}
  \item itkAdaptiveNonLocalMeansDenoisingImageFilter.h
  \item itkAdaptiveNonLocalMeansDenoisingImageFilter.hxx
\end{itemize}
The following helper class
\begin{itemize}
  \item itkVarianceImageFilter.h
  \item itkVarianceImageFilter.hxx
\end{itemize}
is the variance analog to the \doxygen{MeanImageFilter} which ``Computes an image where a
given pixel is the [variance] value of the the pixels in a neighborhood about the corresponding
input pixel.''  Both the adaptive and variance filters are multi-threaded.

Additionally, we wrote an easy-to-use interface program, DenoiseImage.cxx, with an ANTs-based command
line interface facilitated by the following helper classes:
\begin{itemize}
  \item antsCommandLineOption.h
  \item antsCommandLineOption.cxx
  \item antsCommandLineParser.h
  \item antsCommandLineParser.cxx
\end{itemize}
These latter files are not essential for other applications or for inclusion into ITK.
They are included simply to facilitate the generation of a stand-alone application that
people can use.  One can invoke the help menu via \verb#$ DenoiseImage --help 1# at the
command line:
\small
\begin{verbatim}
$ DenoiseImage --help 1

COMMAND:
     DenoiseImage
          Denoise an image using a spatially adaptive filter originally described in J. V.
          Manjon, P. Coupe, Luis Marti-Bonmati, D. L. Collins, and M. Robles. Adaptive
          Non-Local Means Denoising of MR Images With Spatially Varying Noise Levels,
          Journal of Magnetic Resonance Imaging, 31:192-203, June 2010.

OPTIONS:
     -d, --image-dimensionality 2/3/4
          This option forces the image to be treated as a specified-dimensional image. If
          not specified, the program tries to infer the dimensionality from the input
          image.

     -i, --input-image inputImageFilename
          A scalar image is expected as input for noise correction.

     -n, --noise-model Rician/(Gaussian)
          Employ a Rician or Gaussian noise model.

     -x, --mask-image maskImageFilename
          If a mask image is specified, denoising is only performed in the mask region.

     -s, --shrink-factor (1)/2/3/...
          Running noise correction on large images can be time consuming. To lessen
          computation time, the input image can be resampled. The shrink factor, specified
          as a single integer, describes this resampling. Shrink factor = 1 is the
          default.

     -p, --patch-radius 1
                        1x1x1
          Patch radius. Default = 1x1x1

     -r, --search-radius 3
                         3x3x3
          Search radius. Default = 3x3x3.

     -o, --output correctedImage
                  [correctedImage,<noiseImage>]
          The output consists of the noise corrected version of the input image.
          Optionally, one can also output the estimated noise image.

     -v, --verbose (0)/1
          Verbose output.

     -h
          Print the help menu (short version).

     --help
          Print the help menu.
\end{verbatim}
\normalsize
which should provide minimal detail regarding each parameter.  The default parameters
should provide adequate performance for a variety of cases thus requiring only the simple
call (for a 3-D image)

\small
\begin{verbatim}
$ DenoiseImage -d 3 -i ${inputImage} -o ${outputImage}
\end{verbatim}
\normalsize

As mentioned, this filter is multi-threaded so one might want to set the environment
variable to enable more than one thread, e.g.,
\small
\begin{verbatim}
$ export ITK_GLOBAL_DEFAULT_NUMBER_OF_THREADS=4
\end{verbatim}
\normalsize

To see the progress of the filter, one also needs to set the verbose flag, i.e.,
\verb#-v 1#.

\section{Examples}

We included a 2-D brain slice and a 3-D hyperpolarized gas volume each with a corresponding
mask.  Below we include the call and resulting output for these two sets of data.

\subsection{T1-weighted MRI brain slice}

\small
\begin{verbatim}
$ DenoiseImage -d 2 -i t81slice.nii.gz -x t81mask.nii.gz -o t81denoised.nii.gz -v 1

Running for 2-dimensional images.

*********10*********20*********30*********40*********50*********60*********70*********80*********90*********

   AdaptiveNonLocalMeansDenoisingImageFilter (0x7f9733997080)
     RTTI typeinfo:   itk::AdaptiveNonLocalMeansDenoisingImageFilter<itk::Image<float, 2u>, itk::Image<float, 2u>, itk::Image<unsigned char, 2u> >
     Reference Count: 1
     Modified Time: 517
     Debug: Off
     Object Name:
     Observers:
       ProgressEvent(Command)
     Inputs:
       Primary: (0x7f9733996c10) *
       _1: (0x7f97339988f0)
     Indexed Inputs:
       0: Primary (0x7f9733996c10)
       1: _1 (0x7f97339988f0)
     Required Input Names: Primary
     NumberOfRequiredInputs: 1
     Outputs:
       Primary: (0x7f9733996a50)
     Indexed Outputs:
       0: Primary (0x7f9733996a50)
     NumberOfRequiredOutputs: 1
     Number Of Threads: 1
     ReleaseDataFlag: Off
     ReleaseDataBeforeUpdateFlag: Off
     AbortGenerateData: Off
     Progress: 1
     Multithreader:
       RTTI typeinfo:   itk::MultiThreader
       Reference Count: 1
       Modified Time: 299
       Debug: Off
       Object Name:
       Observers:
         none
       Thread Count: 1
       Global Maximum Number Of Threads: 128
       Global Default Number Of Threads: 1
     CoordinateTolerance: 1e-06
     DirectionTolerance: 1e-06
     Using Gaussian noise model.
     Epsilon = 1e-05
     Mean threshold = 0.95
     Variance threshold = 0.5
     Smoothing variance = 2
     Neighborhood radius for local mean and variance = [1, 1]
     Neighborhood search radius  = [3, 3]
     Neighborhood block radius = [1, 1]
Elapsed time: 0.0831451
\end{verbatim}
\normalsize

\begin{figure}[!ht]
\begin{subfigure}{0.5\textwidth}
  \centering
  \includegraphics[width=0.8\linewidth]{Figures/t81slice.eps}
  \caption{Original}
\end{subfigure}
\begin{subfigure}{0.5\textwidth}
  \centering
  \includegraphics[width=0.8\linewidth]{Figures/t81denoised.eps}
  \caption{Denoised}
\end{subfigure}
\caption{Effects of the NL-means filter on a 2-D T1-weighted MRI brain slice.}
\end{figure}

\subsection{Hyperpolarized He3 Lung MRI volume}

\small
\begin{verbatim}
$ DenoiseImage -d 3 -i he3volume.nii.gz -x he3mask.nii.gz -o he3denoised.nii.gz -v 1

Running for 3-dimensional images.

*********10*********20*********30*********40*********50*********60*********70*********80*********90*********

   AdaptiveNonLocalMeansDenoisingImageFilter (0x7ff279e0a440)
     RTTI typeinfo:   itk::AdaptiveNonLocalMeansDenoisingImageFilter<itk::Image<float, 3u>, itk::Image<float, 3u>, itk::Image<unsigned char, 3u> >
     Reference Count: 1
     Modified Time: 506
     Debug: Off
     Object Name:
     Observers:
       ProgressEvent(Command)
     Inputs:
       Primary: (0x7ff27a97ea20) *
       _1: (0x7ff27a97f040)
     Indexed Inputs:
       0: Primary (0x7ff27a97ea20)
       1: _1 (0x7ff27a97f040)
     Required Input Names: Primary
     NumberOfRequiredInputs: 1
     Outputs:
       Primary: (0x7ff279e0a710)
     Indexed Outputs:
       0: Primary (0x7ff279e0a710)
     NumberOfRequiredOutputs: 1
     Number Of Threads: 4
     ReleaseDataFlag: Off
     ReleaseDataBeforeUpdateFlag: Off
     AbortGenerateData: Off
     Progress: 1
     Multithreader:
       RTTI typeinfo:   itk::MultiThreader
       Reference Count: 1
       Modified Time: 294
       Debug: Off
       Object Name:
       Observers:
         none
       Thread Count: 4
       Global Maximum Number Of Threads: 128
       Global Default Number Of Threads: 4
     CoordinateTolerance: 1e-06
     DirectionTolerance: 1e-06
     Using Gaussian noise model.
     Epsilon = 1e-05
     Mean threshold = 0.95
     Variance threshold = 0.5
     Smoothing variance = 2
     Neighborhood radius for local mean and variance = [1, 1, 1]
     Neighborhood search radius  = [3, 3, 3]
     Neighborhood block radius = [1, 1, 1]
Elapsed time: 1.7134
\end{verbatim}
\normalsize

\begin{figure}[!ht]
\begin{subfigure}{0.5\textwidth}
  \centering
  \includegraphics[width=0.8\linewidth]{Figures/he3volume.eps}
  \caption{Original}
\end{subfigure}
\begin{subfigure}{0.5\textwidth}
  \centering
  \includegraphics[width=0.8\linewidth]{Figures/he3denoised.eps}
  \caption{Denoised}
\end{subfigure}
\caption{Effects of the NL-means filter on a 3-D hyperpolarized gas MRI lung volume.}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  Insert the bibliography using BibTeX
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliographystyle{plain}
\bibliography{InsightJournal}

\end{document}