### 2017-03-15 Document/LaTeX/ArticleExample.tex
ENH:  UPdate text.

%
% Complete documentation on the extended LaTeX markup used for Insight
% documentation is available in ``Documenting Insight'', which is part
% of the standard documentation for Insight.  It may be found online
% at:
%
%     http://www.itk.org/

\documentclass{InsightArticle}

\usepackage[dvips]{graphicx}
\usepackage[cmtip,all]{xy}
\newcommand{\longsquiggly}{\xymatrix{{}\ar@{~>}[r]&{}}}
\usepackage{mathtools}
\usepackage{subcaption}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  hyperref should be the last package to be loaded.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage[
bookmarks,
bookmarksopen,
backref,
colorlinks,linkcolor={blue},citecolor={blue},urlcolor={blue},
]{hyperref}

%  This is a template for Papers to the Insight Journal.
%  It is comparable to a technical report format.

% The title should be descriptive enough for people to be able to find
% the relevant document.
\title{A patch-based framework for new ITK functionality:  Joint fusion, denoising,
and non-local super-resolution}

%
% NOTE: This is the last number of the "handle" URL that
% The Insight Journal assigns to your paper as part of the
% submission process. Please replace the number "1338" with
% the actual handle number that you get assigned.
%
\newcommand{\IJhandlerIDnumber}{}

% Increment the release number whenever significant changes are made.
% The author and/or editor can define 'significant' however they like.
\release{0.00}

% At minimum, give your name and an email address.  You can include a
% snail-mail address if you like.
\author{Nick Tustison$^{1}$ Brian Avants$^{2}$, Hongzhi Wang$^{3}$, Long Xie$^{4}$, Paul Yushkevich$^{4}$, and Jose Vicente Manj\'on$^{5}$}
\authoraddress{$^{1}$University of Virginia, Charlottesville, VA\\
               $^{2}$Biogen, Cambridge, MA\\
               $^{3}$IBM, San Jose, CA\\
               $^{4}$University of Pennsylviania, Philadelphia, PA\\
               $^{5}$ITACA institute, Universidad Polit\'ecnica de Valencia, Valencia, Spain
               }

\begin{document}

%
% Add hyperlink to the web location and license of the paper.
% The argument of this command is the handler identifier given
% by the Insight Journal to this paper.
%
\IJhandlefooter{\IJhandlerIDnumber}

%\ifpdf
%\else
%   %
%   % Commands for including Graphics when using latex
%   %
%   \DeclareGraphicsExtensions{.eps,.jpg,.gif,.tiff,.bmp,.png}
%   \DeclareGraphicsRule{.jpg}{eps}{.jpg.bb}{`convert #1 eps:-}
%   \DeclareGraphicsRule{.gif}{eps}{.gif.bb}{`convert #1 eps:-}
%   \DeclareGraphicsRule{.tiff}{eps}{.tiff.bb}{`convert #1 eps:-}
%   \DeclareGraphicsRule{.bmp}{eps}{.bmp.bb}{`convert #1 eps:-}
%   \DeclareGraphicsRule{.png}{eps}{.png.bb}{`convert #1 eps:-}
%\fi

\maketitle

\ifhtml
\chapter*{Front Matter\label{front}}
\fi

% The abstract should be a paragraph or two long, and describe the
% scope of the document.
\begin{abstract}
\noindent
In an earlier Insight Journal article, we introduced an ITK implementation of the
adaptive patch-based image denoising algorithm described in \cite{Manjon:2010aa}.
We follow-up up that offering with a generalized non-local, patch-based ITK class
framework and a refactored denoising class.  In addition, we provide two
ITK implementations of related, well-known algorithms.  The first is a non-local
super resolution method described in \cite{Manjon:2010ab,Manjon:2010ac}.  The second is
the multivariate
joint label fusion algorithm of \cite{Wang:2013ab,Wang:2013aa} with additional extensions,
denoted as ``joint intensity fusion'', which will be described in a forthcoming manuscript.
Accompanying these ITK classes are
documented programming interfaces which use our previously introduced unique
command line interface routines.  Several 2-D examples on brain imaging data are
provided to qualitatively demonstrate performance.
\end{abstract}

\IJhandlenote{\IJhandlerIDnumber}

% \tableofcontents

\section{Introduction}

Since the non-local (NL) means algorithm introduced in several early papers \cite{Buades},
the resulting family of research techniques has grown to significant prominence within the
medical imaging community.  In addition to continued research output, patch-based techniques
have been the subject of workshops (Patch-MI 2015 and PMI'16, both MICCAI-sponsored
workshops).   Given the interest and potential applications, this submission is a potential
``conversation starter'' for the development of an ITK software framework

\section{ITK Implementation}

The following files define the adaptive NL-means algorithm and is written in ITK
coding style:
\begin{itemize}
  \item itkAdaptiveNonLocalMeansDenoisingImageFilter.h
  \item itkAdaptiveNonLocalMeansDenoisingImageFilter.hxx
\end{itemize}
The following helper class
\begin{itemize}
  \item itkVarianceImageFilter.h
  \item itkVarianceImageFilter.hxx
\end{itemize}
is the variance analog to the \doxygen{MeanImageFilter} which ``Computes an image where a
given pixel is the [variance] value of the the pixels in a neighborhood about the corresponding
input pixel.''  Both the adaptive and variance filters are multi-threaded.

Additionally, we wrote an easy-to-use interface program, DenoiseImage.cxx, with an ANTs-based command
line interface facilitated by the following helper classes:
\begin{itemize}
  \item antsCommandLineOption.h
  \item antsCommandLineOption.cxx
  \item antsCommandLineParser.h
  \item antsCommandLineParser.cxx
\end{itemize}
These latter files are not essential for other applications or for inclusion into ITK.
They are included simply to facilitate the generation of a stand-alone application that
people can use.  One can invoke the help menu via \verb#$ DenoiseImage --help 1# at the
command line:
\small
\begin{verbatim}
$ DenoiseImage --help 1

COMMAND:
     DenoiseImage
          Denoise an image using a spatially adaptive filter originally described in J. V.
          Manjon, P. Coupe, Luis Marti-Bonmati, D. L. Collins, and M. Robles. Adaptive
          Non-Local Means Denoising of MR Images With Spatially Varying Noise Levels,
          Journal of Magnetic Resonance Imaging, 31:192-203, June 2010.

OPTIONS:
     -d, --image-dimensionality 2/3/4
          This option forces the image to be treated as a specified-dimensional image. If
          not specified, the program tries to infer the dimensionality from the input
          image.

     -i, --input-image inputImageFilename
          A scalar image is expected as input for noise correction.

     -n, --noise-model Rician/(Gaussian)
          Employ a Rician or Gaussian noise model.

     -x, --mask-image maskImageFilename
          If a mask image is specified, denoising is only performed in the mask region.

     -s, --shrink-factor (1)/2/3/...
          Running noise correction on large images can be time consuming. To lessen
          computation time, the input image can be resampled. The shrink factor, specified
          as a single integer, describes this resampling. Shrink factor = 1 is the
          default.

     -p, --patch-radius 1
                        1x1x1
          Patch radius. Default = 1x1x1

     -r, --search-radius 3
                         3x3x3
          Search radius. Default = 3x3x3.

     -o, --output correctedImage
                  [correctedImage,<noiseImage>]
          The output consists of the noise corrected version of the input image.
          Optionally, one can also output the estimated noise image.

     -v, --verbose (0)/1
          Verbose output.

     -h
          Print the help menu (short version).

     --help
          Print the help menu.
\end{verbatim}
\normalsize
which should provide minimal detail regarding each parameter.  The default parameters
should provide adequate performance for a variety of cases thus requiring only the simple
call (for a 3-D image)

\small
\begin{verbatim}
$ DenoiseImage -d 3 -i ${inputImage} -o ${outputImage}
\end{verbatim}
\normalsize

As mentioned, this filter is multi-threaded so one might want to set the environment
variable to enable more than one thread, e.g.,
\small
\begin{verbatim}
$ export ITK_GLOBAL_DEFAULT_NUMBER_OF_THREADS=4
\end{verbatim}
\normalsize

To see the progress of the filter, one also needs to set the verbose flag, i.e.,
\verb#-v 1#.

\section{Examples}

We included a 2-D brain slice and a 3-D hyperpolarized gas volume each with a corresponding
mask.  Below we include the call and resulting output for these two sets of data.

\subsection{T1-weighted MRI brain slice}

\small
\begin{verbatim}
$ DenoiseImage -d 2 -i t81slice.nii.gz -x t81mask.nii.gz -o t81denoised.nii.gz -v 1

Running for 2-dimensional images.

*********10*********20*********30*********40*********50*********60*********70*********80*********90*********

   AdaptiveNonLocalMeansDenoisingImageFilter (0x7f9733997080)
     RTTI typeinfo:   itk::AdaptiveNonLocalMeansDenoisingImageFilter<itk::Image<float, 2u>, itk::Image<float, 2u>, itk::Image<unsigned char, 2u> >
     Reference Count: 1
     Modified Time: 517
     Debug: Off
     Object Name:
     Observers:
       ProgressEvent(Command)
     Inputs:
       Primary: (0x7f9733996c10) *
       _1: (0x7f97339988f0)
     Indexed Inputs:
       0: Primary (0x7f9733996c10)
       1: _1 (0x7f97339988f0)
     Required Input Names: Primary
     NumberOfRequiredInputs: 1
     Outputs:
       Primary: (0x7f9733996a50)
     Indexed Outputs:
       0: Primary (0x7f9733996a50)
     NumberOfRequiredOutputs: 1
     Number Of Threads: 1
     ReleaseDataFlag: Off
     ReleaseDataBeforeUpdateFlag: Off
     AbortGenerateData: Off
     Progress: 1
     Multithreader:
       RTTI typeinfo:   itk::MultiThreader
       Reference Count: 1
       Modified Time: 299
       Debug: Off
       Object Name:
       Observers:
         none
       Thread Count: 1
       Global Maximum Number Of Threads: 128
       Global Default Number Of Threads: 1
     CoordinateTolerance: 1e-06
     DirectionTolerance: 1e-06
     Using Gaussian noise model.
     Epsilon = 1e-05
     Mean threshold = 0.95
     Variance threshold = 0.5
     Smoothing variance = 2
     Neighborhood radius for local mean and variance = [1, 1]
     Neighborhood search radius  = [3, 3]
     Neighborhood block radius = [1, 1]
Elapsed time: 0.0831451
\end{verbatim}
\normalsize

\begin{figure}[!ht]
\begin{subfigure}{0.5\textwidth}
  \centering
  \includegraphics[width=0.8\linewidth]{Figures/t81slice.eps}
  \caption{Original}
\end{subfigure}
\begin{subfigure}{0.5\textwidth}
  \centering
  \includegraphics[width=0.8\linewidth]{Figures/t81denoised.eps}
  \caption{Denoised}
\end{subfigure}
\caption{Effects of the NL-means filter on a 2-D T1-weighted MRI brain slice.}
\end{figure}

\subsection{Hyperpolarized He3 Lung MRI volume}

\small
\begin{verbatim}
$ DenoiseImage -d 3 -i he3volume.nii.gz -x he3mask.nii.gz -o he3denoised.nii.gz -v 1

Running for 3-dimensional images.

*********10*********20*********30*********40*********50*********60*********70*********80*********90*********

   AdaptiveNonLocalMeansDenoisingImageFilter (0x7ff279e0a440)
     RTTI typeinfo:   itk::AdaptiveNonLocalMeansDenoisingImageFilter<itk::Image<float, 3u>, itk::Image<float, 3u>, itk::Image<unsigned char, 3u> >
     Reference Count: 1
     Modified Time: 506
     Debug: Off
     Object Name:
     Observers:
       ProgressEvent(Command)
     Inputs:
       Primary: (0x7ff27a97ea20) *
       _1: (0x7ff27a97f040)
     Indexed Inputs:
       0: Primary (0x7ff27a97ea20)
       1: _1 (0x7ff27a97f040)
     Required Input Names: Primary
     NumberOfRequiredInputs: 1
     Outputs:
       Primary: (0x7ff279e0a710)
     Indexed Outputs:
       0: Primary (0x7ff279e0a710)
     NumberOfRequiredOutputs: 1
     Number Of Threads: 4
     ReleaseDataFlag: Off
     ReleaseDataBeforeUpdateFlag: Off
     AbortGenerateData: Off
     Progress: 1
     Multithreader:
       RTTI typeinfo:   itk::MultiThreader
       Reference Count: 1
       Modified Time: 294
       Debug: Off
       Object Name:
       Observers:
         none
       Thread Count: 4
       Global Maximum Number Of Threads: 128
       Global Default Number Of Threads: 4
     CoordinateTolerance: 1e-06
     DirectionTolerance: 1e-06
     Using Gaussian noise model.
     Epsilon = 1e-05
     Mean threshold = 0.95
     Variance threshold = 0.5
     Smoothing variance = 2
     Neighborhood radius for local mean and variance = [1, 1, 1]
     Neighborhood search radius  = [3, 3, 3]
     Neighborhood block radius = [1, 1, 1]
Elapsed time: 1.7134
\end{verbatim}
\normalsize

\begin{figure}[!ht]
\begin{subfigure}{0.5\textwidth}
  \centering
  \includegraphics[width=0.8\linewidth]{Figures/he3volume.eps}
  \caption{Original}
\end{subfigure}
\begin{subfigure}{0.5\textwidth}
  \centering
  \includegraphics[width=0.8\linewidth]{Figures/he3denoised.eps}
  \caption{Denoised}
\end{subfigure}
\caption{Effects of the NL-means filter on a 3-D hyperpolarized gas MRI lung volume.}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  Insert the bibliography using BibTeX
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliographystyle{plain}
\bibliography{InsightJournal}

\end{document}